---
title: "a2_task2_mariano_viz"
author: "Mariano Viz"
date: "06/02/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(patchwork)
library(broom)
library(jtools)
library(kableExtra)

```

### Summary
In this report, I explore the abundance of the mountain yellow-legged frog (Rana muscosa) in the Sierra Nevada ecosystem for the period 1995-2002. Specifically, I examine the total number of adult, subadult, and tadpole observed in the entire study region by year and the total counts of combined adult and subadult observed in the 5 lakes with the greatest total counts. Results are presented in a single compound figure. Data were extracted from Knapp et al. (2020).



```{r}
#Read in data
palmetto_data <- read_csv(here("data", "palmetto.csv"))
  

```



```{r}
#Data wrangling:
palmetto_subset <- palmetto_data %>% 
  select(species, height, length, width, green_lvs) %>% #selecting relevant columns
  mutate(species = case_when(species == "1" ~ "Serenoa repens", species == "2" ~ "Sabal etonia")) #renaming species
```


### Data Visualization

```{r}

##Data visualization (stored to make a compound figure): 

#Data visualization (Species and Height):
sp_height_plot <- ggplot(data = palmetto_subset, aes(x = species, y = height)) +
  geom_violin(aes(color = species, fill=species), alpha = 0.2)+
  geom_boxplot(fill = NA, 
               width = 0.2, 
               outlier.color = NA)+
  stat_summary(fun = mean,
               alpha = 0.6,
               size = 0.3) +
  scale_fill_manual(values = c("orange2","palegreen4")) +
  scale_color_manual(values = c("orange2","palegreen4"))+
  labs(x = "Species",
       y = "Height (cm)") +
  theme_minimal() +
  theme(legend.position="none")

#Data visualization (Species and Canopy length):
sp_length_plot <- ggplot(data = palmetto_subset, aes(x = species, y = length)) +
  geom_violin(aes(color = species, fill=species), alpha = 0.2)+
  geom_boxplot(fill = NA, 
               width = 0.2, 
               outlier.color = NA)+
  stat_summary(fun = mean,
               alpha = 0.6,
               size = 0.3) +
  scale_fill_manual(values = c("orange2","palegreen4")) +
  scale_color_manual(values = c("orange2","palegreen4"))+
  labs(x = "Species",
       y = "Canopy length (cm)") +
  theme_minimal() +
  theme(legend.position="none")
  

#Data visualization stored for (Species and Canopy width):
sp_width_plot <- ggplot(data = palmetto_subset, aes(x = species, y = width)) +
  geom_violin(aes(color = species, fill=species), alpha = 0.2)+
  geom_boxplot(fill = NA, 
               width = 0.2, 
               outlier.color = NA)+
  stat_summary(fun = mean,
               alpha = 0.6,
               size = 0.3) +
  scale_fill_manual(values = c("orange2","palegreen4")) +
  scale_color_manual(values = c("orange2","palegreen4"))+
  labs(x = "Species",
       y = "Canopy width (cm)") +
  theme_minimal() +
  theme(legend.position="none")


#Compound figure (patchwork):
sp_height_plot | sp_length_plot | sp_width_plot 
  

```

**Figure 1.** Data: Abrahamson, W.G. (2019).



```{r}
#Data visualization (Canopy length and Canopy width for both species):
ggplot(data = palmetto_subset, aes(x = length, y = width)) +
  geom_point(aes(color = species), alpha = 0.2) +
  scale_color_manual(values = c("orange2","palegreen4"))+
  labs(x = "Canopy length (cm)",
       y = "Canopy width (cm)") +
  theme_minimal() 
```

**Figure 2.** Data: Abrahamson, W.G. (2019).


```{r}
#Data visualization (Species and Count of green leaves):
ggplot(data = palmetto_subset, aes(x = species, y = green_lvs)) +
  geom_violin(aes(color = species, fill=species), alpha = 0.2)+
  geom_boxplot(fill = NA, 
               width = 0.2, 
               outlier.color = NA)+
  stat_summary(fun = mean,
               alpha = 0.6,
               size = 0.3) +
  scale_fill_manual(values = c("orange2","palegreen4")) +
  scale_color_manual(values = c("orange2","palegreen4"))+
  labs(x = "Species",
       y = "Canopy Width (cm)") +
  theme_minimal() +
  theme(legend.position="none")

```

**Figure 3.** Data: Abrahamson, W.G. (2019).




### Binary Logistic Regression


**Table 1.** Binary Logistic Regression model rersults. Data: Abrahamson, W.G. (2019).
```{r}
#Convert Species column to factor:
palmetto_blr <- palmetto_subset %>% 
mutate(species = as.factor(species))

#Binary Logistic Regression:
palmetto_species_blr <- glm(species ~ height + length + width + green_lvs, 
                            data = palmetto_blr,
                            family = "binomial")

#Observation of the outcomes:
    #palmetto_species_blr 
    #summary(palmetto_species_blr)


#Finalized table with model outcomes:
palmetto_species_blr_tidy <-
  broom::tidy(palmetto_species_blr) %>% 
  select(term, estimate, std.error, p.value) %>%
  mutate(p.value = case_when(p.value != "1" ~ "<2e-16")) %>% 
  kable(col.names = c("Variable", "Coefficient", "Standar Error", "Significance (p.value)")) %>% 
  kable_styling(bootstrap_options = "striped")

palmetto_species_blr_tidy


```




### Model's capabilities to correctly classify species (Serenoa repens / Sabal etonia)

**Table 2.** Binary Logistic Regression model rersults. Data: Abrahamson, W.G. (2019).
```{r}
#Probability of being Serenoa repens (non-zero level):
blr_fitted <- palmetto_species_blr %>% 
  broom::augment(type.predict = "response")



#Table about how successfully this model classify a plant as the correct species: 
blr_prob <- blr_fitted %>% 
  select(species, .fitted) %>% 
  mutate(classification = case_when(.fitted >= "0.5" ~ "Serenoa repens", 
                                    TRUE ~ "Sabal etonia")) %>%
  mutate(prediction = case_when(species == classification ~ "Correct",
                             species != classification ~ "Incorrect")) %>%
  group_by(species, prediction) %>% 
  summarise(n()) %>% 
  pivot_wider(names_from = prediction, values_from = "n()") %>% 
  mutate(percent_correct = round(100*(Correct/(Correct+Incorrect)), digits=2)) %>% 
  kable(col.names = c("Species", "Correct", "Incrorrect", "Percent correctly classified")) %>% 
  kable_styling(bootstrap_options = "striped")



blr_prob

```








Data source: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5